The Talking Heads Experiment

### 2015-04-27 ind-steelsbook.tex
changes to main tex file
updated labels, image widths, quotation marks, capitalisation, etc. for chapters 1-7
minor changes
hyphenation, orphans, widows and some other layout updated
uncommented all lines from master tex file
fixed some more widows/orphans
hyphenations, footnotes, ...
corrected one too large float
fixed index entries and extra papers, added proofreaders and typesetters

\documentclass[ number=1
			   ,series=cmle
			   ,url=http://langsci-press.org/ 
			   ,isbn=978-3-944675-42-8
			   ,output=short   % long|short|inprep              
			   %,blackandwhite
			   ,smallfont
			   ,biblatex
			   ,draftmode  
			   ,draft
			  ]{LSP/langsci}                          
\usepackage{LSP/lsp-styles/lsp-gb4e}		% verhindert Komma bei mehrfachen Fußnoten?
\usepackage{LSP/lsp-styles/avm}                                                    
\usepackage{layout}  
\usepackage{lipsum}   
% \usepackage{covington}
% \usepackage{tipa}
% \usepackage[T1]{fontenc}
% \usepackage{tipx}
\usepackage{qtree}
\usepackage{amsmath}
\usepackage{multirow} 
\usepackage{multicol}
%\usepackage[latin1]{inputenc}
\usepackage{authorindex}
%\usepackage[round]{natbib}

\usepackage{localcommands}

\title{The Talking Heads Experiment}
\subtitle{Origins of Words and Meanings}
\BackTitle{The Talking Heads Experiment}% - \\ Origins of Words and Meanings}
\author{Luc Steels}
%\Affiliation{ Catalan Institute for Advanced Studies (ICREA) \\ %- Instituci\accute{o} Catalana de Recerca i Estudis Avan\cats\\
%Institut de Biologia Evolutiva (Universitat Pompeu Fabra and CSIC) Barcelona}
\dedication{For my daughter Lenie}
\BackBody{The Talking Heads Experiment, conducted in the years 1999--2001, was the first large-scale experiment 
in which open populations of situated embodied agents created for the first time ever a new shared vocabulary
by playing language games about real world scenes in front of them. The agents could teleport to different 
physical sites in the world through the Internet. Sites, in Antwerp, Brussels, Paris, Tokyo, London, Cambridge
and several other locations were linked into the network. Humans could interact with the robotic agents either
on site or remotely through the Internet and thus influence the evolving ontologies and languages of the 
artificial agents.

@@

\part{The 1999 Talking Heads Book}

%\thispagestyle{empty}
\begin{figure}[htbp]
  \centerline{\includegraphics[width=1.0\textwidth]{extra/figs/cover-original-book}}
\caption{\small\label{left-right} Original cover of the Talking Heads book published in 1999 on the occasion of the 
Laboratorium exhibition in Antwerp.}
\label{f:cover}
\end{figure}

\include{chap1/chap1}
\include{chap2/chap2}
\include{chap3/chap3}
\include{chap4/chap4}
\include{chap5/chap5}
\include{chap6/chap6}
\include{chap7/chap7}

\part{Installations and experiments}
\include{chap8/chap8}
\include{chap9/chap9}
% 
\part{Beyond the Talking Heads}

\include{chap10/chap10}
\include{chap11/chap11}
\include{chap12/chap12}

% \part{Appendices}

% \include{extra/papers-language-steels}

% \listoffigures
% 
% \listoftables

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                                              %%%
%%%             Backmatter                       %%%
%%%                                              %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\backmatter
\phantomsection%this allows hyperlink in ToC to work
\addcontentsline{toc}{chapter}{List of references}
\printbibliography %change to the name of your bib file
\clearpage

@@

\phantomsection%this allows hyperlink in ToC to work
\addcontentsline{toc}{section}{Language index}
\ohead{Language index} 
\printindex[lan]

>>>>

\documentclass[ number=1
			   ,series=cmle
			   ,url=http://langsci-press.org/catalog/book/49 
			   ,isbn=978-3-944675-42-8
			   ,output=short   % long|short|inprep              
			   %,blackandwhite
			   ,smallfont
			   ,biblatex
% 			   ,draftmode  
			   %,draft
			  ]{LSP/langsci}                          
\usepackage{LSP/lsp-styles/lsp-gb4e}		% verhindert Komma bei mehrfachen Fußnoten?
\usepackage{LSP/lsp-styles/avm}  
\usepackage{todonotes}                                                  
\usepackage{layout}  
\usepackage{lipsum}   
% \usepackage{covington}
% \usepackage{tipa}
% \usepackage[T1]{fontenc}
% \usepackage{tipx}
\usepackage{qtree}
\usepackage{amsmath}
\usepackage{multirow} 
\usepackage{multicol}
%\usepackage[latin1]{inputenc}
\usepackage{authorindex}
%\usepackage[round]{natbib}

\usepackage{localcommands}
\usepackage{localhyphenation}

\proofreader{Martin Hilpert, Felix Kopecky, Peter Petré, Daniela Schröder}
\typesetter{Felix Kopecky, Sebastian Nordhoff}
\title{The Talking Heads Experiment}
\subtitle{Origins of words and meanings}
\BackTitle{The Talking Heads Experiment}% - \\ Origins of Words and Meanings}
\author{Luc Steels}
%\Affiliation{ Catalan Institute for Advanced Studies (ICREA) \\ %- Instituci\accute{o} Catalana de Recerca i Estudis Avan\cats\\
%Institut de Biologia Evolutiva (Universitat Pompeu Fabra and CSIC) Barcelona}
\dedication{For my daughter Lenie}
\BackBody{The Talking Heads Experiment, conducted in the years 1999--2001, was the first large-scale experiment 
in which open populations of situated embodied agents created for the first time ever a new shared vocabulary
by playing language games about real world scenes in front of them. The agents could teleport to different 
physical sites in the world through the Internet. Sites, in Antwerp, Brussels, Paris, Tokyo, London, Cambridge
and several other locations were linked into the network. Humans could interact with the robotic agents either
on site or remotely through the Internet and thus influence the evolving ontologies and languages of the 
artificial agents.

@@

\part{The 1999 Talking Heads book} \label{part:1}

%\thispagestyle{empty}
\begin{figure}[htbp]
  \centerline{\includegraphics[width=.95\textwidth]{extra/figs/cover-original-book.pdf}}
\caption{\small\label{left-right} Original cover of the Talking Heads book published in 1999 on the occasion of the 
Laboratorium exhibition in Antwerp.}
\label{f:cover}
\end{figure}
%
\include{chap1/chap1}
\include{chap2/chap2}
\include{chap3/chap3}
\include{chap4/chap4}
\include{chap5/chap5}
\include{chap6/chap6}
\include{chap7/chap7}
%
\part{Installations and experiments} \label{part:2}
\include{chap8/chap8}
\include{chap9/chap9}

\part{Beyond the Talking Heads} \label{part:3}
\include{chap10/chap10}
\include{chap11/chap11}
\include{chap12/chap12}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                                              %%%
%%%             Backmatter                       %%%
%%%                                              %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\backmatter

% \part{Appendices} \label{part:4}
\include{extra/papers-language-steels}

% \phantomsection%this allows hyperlink in ToC to work
% \addcontentsline{toc}{chapter}{List of references}
\printbibliography %change to the name of your bib file
\clearpage

@@

% \phantomsection%this allows hyperlink in ToC to work
% \addcontentsline{toc}{section}{Language index}
% \ohead{Language index} 
% \printindex[lan]

### 2015-04-27 preface/preface.tex
fixed index entries and extra papers, added proofreaders and typesetters

Dozens of people worked on the Talking Heads. The initial research started in 1997 at the Artificial Intelligence 
Laboratory of the Free University of Brussels (VUB) funded by a ``Geconcentreerde Onderzoeks Actie''
of the Belgian government. Joris Van Looveren worked closely with me on a first prototype 
that was demonstrated in 1997, using an active vision system custom-built by Tony Belpaeme and segmentation 
algorithms implemented by Danny van Tieghem. Edwin de Jong did the first theoretical investigations of the underlying 
semiotic dynamics. Once the initiative for a public installation was taken,  
the main hub became the Sony Computer Science Laboratory in Paris, 
where I worked together intensely with Fr\'{e}d\'{e}ric Kaplan and Angus McIntyre, with additional contributions for the 
teleportation infrastructure by Silv\`{e}re Tajan and Alexis Agahi. The AI Laboratory of the VUB remained
a second hub where important contributions were 
made most notably by Joris Van Looveren, Tony Belpaeme, Holger Kenn and Mario Campanella. To all of them I 
am grateful that we were able to create such an extremely exciting experiment that stimulated many 
thousands of people to think about language and its origins in new ways.

@@

In 2008/2009 I was a fellow at the Wissenschaftskolleg in Berlin, which was an ideal environment to reflect on our research 
program and how we could proceed. Many new ideas came out of that year, in particular, a realisation that for several of the key
puzzles we were trying to solve, inspiration could be found in evolutionary biology, and this lead to a new path which only now 
is beginning to be explored. It is good to know that after a major collapse in funding around 2009, there is 
again a team of young people assembling to push research on language games further. They include Emilia Garcia Casademont in 
Barcelona, Miquel Cornudella and Paul Van Eecke in Paris, and Yana Knight in Brussels. Financing remains precarious 
but the future is in their hands!

>>>>

Dozens of people worked on the Talking Heads. The initial research started in 1997 at the Artificial Intelligence 
Laboratory of the Free University of Brussels (VUB) funded by a ``Geconcentreerde onderzoeks actie'' (GOA)
of the Belgian government. Joris Van Looveren worked closely with me on a first prototype 
that was demonstrated in 1997, using an active vision system custom-built by Tony Belpaeme and segmentation 
algorithms implemented by Danny van Tieghem. Edwin de Jong did the first theoretical investigations of the underlying 
semiotic dynamics. Once the initiative for a public installation was taken,  
the main hub became the Sony Computer Science Laboratory in Paris, 
where I worked together intensely with Fr\'{e}d\'{e}ric Kaplan and Angus McIntyre, with additional contributions for the 
teleportation infrastructure by Silv\`{e}re Tajan and Alexis Agahi. The AI Laboratory of the VUB remained
a second hub where important contributions were 
made most notably by Joris Van Looveren, Tony Belpaeme, Holger Kenn and Mario Campanella. To all of them I 
am grateful that we were able to create such an extremely exciting experiment that stimulated many 
thousands of people to think about language and its origins in new ways.

@@

In 2008/2009 I was a fellow at the Wissenschaftskolleg in Berlin, which was an ideal environment to reflect on our research 
program and how we could proceed. Many new ideas came out of that year, in particular, a realisation that for several of the key
puzzles we were trying to solve, inspiration could be found in evolutionary biology, and this led to a new path which only now 
is beginning to be explored. It is good to know that after a major collapse in funding around 2009, there is 
again a team of young people assembling to push research on language games further. They include Emilia Garcia Casademont in 
Barcelona, Miquel Cornudella and Paul Van Eecke in Paris, and Yana Knight in Brussels. Financing remains precarious 
but the future is in their hands!
